---
title: "R sample of Time Series Analysis"
author: 'Hsin-Yu,Ku (Matrikel Num. : 3038591)'
output:
  pdf_document: 
    keep_tex: yes
    toc_depth: 1
  html_document: default
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.width=5, fig.height=3)
set.seed(1234)
```
goal
    Goal : explain basic concepts of Time Series(TS) Analysis with several TS datasets.
##### What is Time Series (TS) ? 
   TS data is a sequence of data in chronological order.   
   
##### ----------------------------------- 01 / Build TS data from non-TS data ----------------------------------- ######
    ts() turns some data into a TS data.
      Use: ts(data , start=YEAR, frequency=how many frequences are there in a year)  
```{r}
vector_beforeTS <- c(2, 6, 8, 3, 9, 1, 0, 4) 
TS_data <- ts(vector_beforeTS, start=2020, frequency=4)
# print() shows TS_data is a TS data inclding 4 quarters from each year 2020-2021.
print(TS_data)
```
##### --------------------------- 02 / A quick inspection into time Series(TS) data --------------------------- #####
```{r}
?LakeHuron 
# LakeHuron shows annual measurements (in feet) of the level of Lake Huron in 1875–1972.
```
```{r}
# is.ts() checks if a data set is a TS data or not. (TRUE/FALSE)
is.ts(LakeHuron)
```
```{r}
# print() shows that it is a TS data with start in 1875, end in 1972, frequency =1.
print(LakeHuron)
```
```{r}
# str() shows it is a TS data with 98 observations between 1875-1972
str(LakeHuron) 
```
```{r eval=FALSE, include=FALSE}
# To see how many observations are there in a TS data, length() can be used too.
length(LakeHuron) 
```

```{r, out.width = '60%'}
# ts.plot() gives an overlook of the trend of one or more TS in a single common plot. 
ts.plot(LakeHuron, main="Level of Lake Huron(in feet)", ylab=NA)
```

    When there are more than 1 TS trends, use argument "col=" to display with distinct colors.

```{r,fig.width=6, out.width = '70%'}
?EuStockMarkets # EuStockMarkets includes daily price of 4 stock indices (1991-1998).
 # Use col() in ts.plot() to displays trends of 4 stock indices (DAX,SMI,CAC,FTSE) together.
ts.plot(EuStockMarkets,col=1:4, main="Price Trends of 4 Stock Indices")
 # Add a legend on the topleft with 0.7 as size, type of line = 1, width =2.
legend("topleft",colnames(EuStockMarkets), cex=0.7, lty=1, lwd=2, col=1:4)
```

##### -------------------------------- 03/ More functions to look into a TS data -------------------------------- #####

  **start() & end() functions return the time index of the first & last observations.**
```{r}
start(AirPassengers)
end(AirPassengers)
 # AirPassengers is a TS data starts on 1949 with time index 1 & ends on 1960 with time index 12 
```

**deltat() function returns the fixed time interval between observations.**

```{r}
deltat(AirPassengers)
 # AirPassengers has a time interval of 0.083333 year between observations.
```
```{r eval=FALSE, include=FALSE}
# time() function calculates a vector of time indices, with one element for each time index on which the series was observed.**
time(AirPassengers)
```
**frequency() function returns the number of observations per unit time.**
```{r}
frequency(AirPassengers)
 # AirPassengers has 12 observations per unit time(year)
```
**cycle() function returns the position in the cycle of each observation.**
```{r}
cycle(AirPassengers)
```
##### ----------------------------- 04/ Simple transformations of trends in TS data ----------------------------- #####

***Transformation by diff()***
   : diff() deletes seasonal trend in a TS data with "lag = cycle of seasonality." 
```{r,fig1,fig.width=4, fig.height=4,out.width='60%'}
AP_no_cycle <-diff(AirPassengers, lag=12)
plot.ts(cbind(AirPassengers, AP_no_cycle), 
        main= " AirPassengers: with & w/o yearly pattern", cex.main=1, cex.lab=0.9)
# plot.ts() plots several TS on differnt bases (but with same time frequency).
# Here we see that year 1954 and 1958 have comparably low or even negative growth
  # in the overall growing trend, comparing to the same months of the previous year.  
```
***Transformation by log()***
```{r}
?airmiles # airmiles shows miles flown by commercial airlines in the US from 1937 to 1960.
```
```{r,fig.height=4.3, out.width='40%'}
 # log() linarlizes the rapid growth since around 1950. 
plot.ts(cbind(airmiles,log(airmiles)), 
        main="Log Transformation: Before & After", cex.main=1.3)
```

##### ------------------------ 05 / Two TS models : White Noise & Random Walk ------------------------ ####

   ***White Noise (WN) model***
```{r, fig.height=2.8,out.width='50%'}
# Generating WN data with arima.sim(): order=c(0,0,0)
WN_data <- arima.sim(model =list(order=c(0,0,0)), n = 200 )
ts.plot(WN_data, main= "TS data from Whit Noise Model")
```
 
   ***Random Walk (RW) model***
```{r, fig.height=2.8,out.width='50%'}
# Generating RW data with arima.sim(): order=c(0,1,0)
RW_data <- arima.sim(model =list(order=c(0,1,0)), n = 200)
ts.plot(RW_data, main= "TS data from Random Walk Model")
```
```{r,fig.height=2.8, out.width='50%'}
# Use cumsum() to turn WN model data into RW model data.
RW_data2 <- cumsum(WN_data)
ts.plot(RW_data2)
title(main= list("Random Walk data   
generated by cumsum() from White Noise Model",cex=1.2))
```